\bf A Hyper-edge Event Model

### 2018-05-15 paper/paper.tex
added some hyper-edge citations

Several network models have been developed that handle time-stamped events in which edge formation is governed by structural dynamics similar to those used in the ERGM \citep{Butts2008,Vu2011,snijders1996stochastic}. Those models are very useful to understand which traits and behaviours are predictive of interactions, however, none of the models explicitly allows hyper-edges---a connection between two or more vertices generated simultaneously\footnote{A hyper-edge connecting just two nodes is simply a usual dyadic edge.}---which is a common property of digitalized textual interactions such as emails and online messages. For instance, \cite{PerryWolfe2012} treat multicast interactions---one type of directed hyper-edge which involve one sender and multiple receivers---via duplication (i.e., obtain pairwise interactions from the original multicast), but their model parameter estimation relies on likelihood approximation which leads to bias. Similarly, \cite{fan2009learning} treats multicast emails as multiple single-recipient edges and randomly jitter the sent times, to avoid the violation of continous-time model assumption. Concentrating on better handling of hyper-edges, we develop the hyper-edge event model (HeEM) which simultaneously models the two components that govern time-stamped event formation: 1) the recipient selection processs that allows multiple senders or receivers, and 2) the time-to-next interactions with flexible distributional choices.

In what follows, we introduce the HeEM by describing how we assume the generatve process of a time-stamped event data (Section 2), and deriving the sampling equations for Bayesian inference (Section 3). Then, we apply the model to the Montgomery county government email data and perform two model validation tasks (Section 4). Finally, the paper finishes in Section 5 with conclusion and discussion.

>>>>

Several network models have been developed that handle time-stamped events in which edge formation is governed by structural dynamics similar to those used in the ERGM \citep{Butts2008,Vu2011,snijders1996stochastic}. Those models are very useful to understand which traits and behaviours are predictive of interactions, however, none of the models explicitly allows hyper-edges---a connection between two or more vertices generated simultaneously\footnote{A hyper-edge connecting just two nodes is simply a usual dyadic edge.}---which is a common property of digitalized textual interactions such as emails and online messages. For instance, \cite{PerryWolfe2012} treat multicast interactions---one type of directed hyper-edge which involve one sender and multiple receivers---via duplication (i.e., obtain pairwise interactions from the original multicast), but their model parameter estimation relies on likelihood approximation which leads to bias. Similarly, \cite{fan2009learning} treats multicast emails as multiple single-recipient edges and randomly jitter the sent times, to avoid the violation of continous-time model assumption. On the other hand, there have been numerous studies for hyper-edges under the hypergraph theory \citep{karypis1999multilevel} in phyics literature, including a random hypergraph model \citep{ghoshal2009random}, but most of them are designed specifically for engineering applications---e.g., multiple tagging networks \citep{zlatic2009hypergraph,zhang2010hypergraph}, cellular networks \citep{klamt2009hypergraphs}, and personalized recommendations \citep{zhang2010personalized,blattner2009b}. To our knowledge, there is a lack of extensive efforts to incorportate hyper-edge modeling into general statistical framework.

Concentrating on developing a statistical network model with better handling of hyper-edges, we introduce the hyper-edge event model (HeEM) which simultaneously models the two components that govern time-stamped event formation: 1) the recipient selection processs that allows multiple senders or receivers, and 2) the time-to-next interactions with flexible distributional choices. In what follows, we introduce the HeEM by describing how we assume the generatve process of a time-stamped event data (Section 2), and deriving the sampling equations for Bayesian inference (Section 3). Then, we apply the model to the Montgomery county government email data and perform two model validation tasks (Section 4). Finally, the paper finishes in Section 5 with conclusion and discussion.
